JHU Machine Learning - Plotting Predictors

Wages Data Example

```{r, cache=TRUE, warning=FALSE, results='hide'}
library(ISLR); library(ggplot2); library(caret); library(gridExtra);

data(Wage)
summary(Wage)
```

Setting training/testing sets

```{r}
inTrain <- createDataPartition(y=Wage$wage, p=.7, list=FALSE)
training <- Wage[inTrain,]
testing <- Wage[-inTrain,]
dim(testing)
```

Feature Plot
- plots all the features against each other

```{r}
featurePlot(x=training[,c("age", "education", "jobclass")],
            y = training$wage,
            plot="pairs")
```

- top right cell "Y" is the outcome (wage) plotted against age, education, and job class
- here, you want to look for any pattern amongst the top cells; do any of the variables seem to have a relationship with the y variable?
        - there seems to be a trend between education and wage

You can also examine the relationships using ggplot:
- appears to have a slight pattern, though the upper scatter of plots seem to be of interest
```{r, cache=TRUE}
qplot(age, wage, data=training)
```

Looking at the plot again with color by different variables:
- appears that majority of higher earning jobs are in information
- this gives you a way to detect variables that might be important in your model because they show variation in your data
```{r, cache=TRUE}
qplot(age, wage, data=training, color = jobclass)
```

Adding regression smoothers to your data:
- Using smooth function to apply linear smoother 

```{r}
qq <- qplot(age, wage, colour=education, data=training)
qq + geom_smooth(method="lm", formula=y~x)
```

Using cut2 to make factors with the Hmisc package
- Often very useful is to break up variabels into different categories because certain categories seem to have different relationships
        - in the case of this data the wages might have different relationships

```{r}
library(Hmisc)
cutWage <- cut2(training$wage, g=3)
table(cutWage)
```

Using these cuts to then make different kinds of plots:
- looking at wage groups vs age; appears that there is some kind of relationships between wage groups and age
- you might however want to add ontop of the the boxplot the actual points themselves because box plots can often obscure how many points are being shown
        - important to this to see points because one cut may not have enough points to
        for real representation of what's going on
```{r, cache=TRUE}
p1 <- qplot(cutWage, age, data=training, fill=cutWage,
            geom=c("boxplot"))
p2 <- qplot(cutWage, age, data=training, fill=cutWage,
            geom=c("boxplot", "jitter"))
grid.arrange(p1,p2,ncol=2)
```

Using tables:

```{r}
t1 <- table(cutWage, training$jobclass)
t1
```

If you want to get the proportion of a variables' presence:
```{r}
prop.table(t1,1)
```

Density Plots

```{r}
qplot(wage, colour=education, data=training, geom="density")
```

Final Notes:
- When making your plots, only using the training set
        - do NOT use the test set for exploration!
- Things to look for:
        - imbalance in outcomes/predictors
        - outliers
        - groups of points not explained by a predictor
        - Skewed Variables

Links:

* [ggplot2 tutorial](http://rstudio-pubs-static.s3.amazonaws.com/2176_75884214fc524dc0bc2a140573da38bb.html)
* [caret visualizations](http://caret.r-forge.r-project.org/visualizations.html)