JHU Machine Learning - Predicting with regression

Key Ideas
- Fit a simple regression model
- plug in new covarieats and multiply by the coeffecients
- useful when the linear model is (nearly) correct
Pros:
- Easy to implement
- East to interpret
Cons:
- Often poor performance in nonlinear settings

Example: Old Faithful eruptions

```{r}
library(caret); data(faithful); set.seed(333)
inTrain <- createDataPartition(y=faithful$waiting,
                               p = 0.5, list=FALSE)
trainFaith <- faithful[inTrain,]
testFaith <- faithful[-inTrain,]
head(trainFaith)
```

**Plot of the variables**
```{r}
plot(trainFaith$waiting, trainFaith$eruptions, pch=19, col="blue", 
     xlab="Waiting", ylab="Duration")
```

**Fitting a linear model**
```{r}
lm1 <- lm(eruptions ~ waiting, data=trainFaith)
summary(lm1)
```
```{r}
plot(trainFaith$waiting, trainFaith$eruptions, pch=19, col="blue", 
     xlab="Waiting", ylab="Duration")
lines(trainFaith$waiting, lm1$fitted, lwd=3)
```

**Predicting a new value**

```{r}
coef(lm1)[1] + coef(lm1)[2] * 80
```
```{r}
newdata <- data.frame(waiting=80)
predict(lm1, newdata)
```

**Plot predictions - training and test set**
```{r}
par(mfrow = c(1,2))
plot(trainFaith$waiting, trainFaith$eruptions, pch=19, col="blue", 
     xlab="Waiting", ylab="Duration")
lines(trainFaith$waiting, predict(lm1), lwd=3)

plot(testFaith$waiting, testFaith$eruptions, pch=19, col="blue", 
     xlab="Waiting", ylab="Duration")
lines(testFaith$waiting, predict(lm1, newdata=testFaith), lwd=3)

```

Get Training set/Test set errors
```{r}
#Calculate RMSE on training
sqrt(sum((lm1$fitted - trainFaith$eruptions)^2))
```
```{r}
#Calculate RMSE on test
sqrt(sum((predict(lm1, newdata=testFaith) - testFaith$eruptions)^2))
```

- since the test data was not used in creating the algorithm, the RMSE is more realistic here on a new data set as opposed to the training set
- RMSE of test set almost always larger than the training

Prediction Intervals
```{r}
pred1 <- predict(lm1, newdata=testFaith, interval="prediction")
ord <- order(testFaith$waiting)
plot(testFaith$waiting, testFaith$eruptions, pch=19, col="blue")
matlines(testFaith$waiting[ord], pred1[ord,], 
         type = "l",, col=c(1,2,2), lty=c(1,1,1), lwd=3)
```
- the interval captures the region where we expect the predicted values to land
        - shows you the range of possible values that could be predicted as opposed
        to a single prediction
        - can be useful for giving you an idea of how well your model will do on new 
        predictions
        
**Same Process with Caret**
```{r}
modFit <- train(eruptions ~ waiting, data=trainFaith, method="lm")
summary(modFit$finalModel)
```

*Final Notes:*
- Regression models with multiple covariates can be included
- often useful in combination with other models
- links:
* [Elements of statistical learning](http://www-stat.stanford.edu/~tibs/ElemStatLearn/)
* [Modern applied statistics with S](http://www.amazon.com/Modern-Applied-Statistics-W-N-Venables/dp/0387954570)
* [Introduction to statistical learning](http://www-bcf.usc.edu/~gareth/ISL/)
