JHU Machine Learning - Forecasting

Google data example

```{r}
library(quantmod)
from.dat <- as.Date("01/01/08", format="%m/%d/%y")
to.dat <- as.Date("12/31/13", format="%m/%d/%y")
getSymbols("GOOG", src="google", from = from.dat, to = to.dat)
head(GOOG)
```

*Summarize monthly and store as time series*

```{r}
mGoog <- to.monthly(GOOG)
googOpen <- Op(mGoog)
ts1 <- ts(googOpen, frequency=12)
plot(ts1, xlab="Years+1", ylab="GOOG")
```

*Decompose a time series into parts*
- decomposes series into a series of different types of patterns in the data 
```{r}
plot(decompose(ts1), xlab="Years+1")
```

*Building training and test sets*
```{r}
ts1Train <- window(ts1, start=1, end = 5)
ts1Test <- window(ts1, start=5, end = 7-0.01)
ts1Train
```

*Simple moving average*
```{r}
plot(ts1Train)
lines(ma(ts1Train, order=3), col="red")
```

*Could also do exponential smoothing*
```{r}
ets1 <- ets(ts1Train, model="MMM")
fcast <- forecast(ets1)
plot(fcast); lines(ts1Test, col="red")
```

*Get the accuracy*
```{r}
accuracy(facst, ts1Test)
```

*Final Notes*
- Forecasting and timeseries prediction is an entire field
        - Rob Hyndman's [Forecasting: principles and practice](https://www.otexts.org/fpp/) is a good place to start
- Cautions:
        - Be wary of spurious correlations
        - Be careful how far you predict (extrapolation)
        - Be wary of dependencies over time
- See quantmod or quandl packages for finance-related problems

